How to use replication assignments for teaching integrity in empirical archaeology

### 2019-08-15 README.Rmd
finished dealing with reviewers' comments on the ms.

### How to download or install

@@

>>>>

### Overview of contents

The **analysis** directory contains:  
- **paper** has our R Markdown document, paper.Rmd, that contains text and code of our paper. It also has a rendered version, paper.docx, suitable for reading (the code is replaced by figures and tables in this file)    
- **data** has our raw data files    
- **figures** has our plots and other illustrations

### How to run, download or install

The simplest way to explore the text, code and data is to click on [binder](https://mybinder.org/v2/gh/benmarwick/replicationteachingarchaeology/master?urlpath=rstudio) to open an instance of RStudio in your browser, which will have the compendium files ready to work with.

@@

```
remotes::install_github("benmarwick/teachingreplicationinarchaeology")
```

### 2019-08-15 analysis/paper/paper.Rmd
finished dealing with reviewers' comments on the ms.

---
title: "How to use replication assignments for teaching integrity in empirical archaeology"
author:
  - Ben Marwick, bmarwick@uw.edu
  - Liying Wang
  - Ryan Robinson
  - Hope Loiselle
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/society-for-american-archaeology.csl" # Insert path for the bib-style
abstract: |
  The value of new archaeological knowledge is strongly determined by how credible it is, and a key measure of scientific credibility is how replicable new results are. However, few archaeologists learn the skills necessary to conduct replication as part of their training. This means there is a gap between the ideals of archaeological science, and the skills we teach future researchers. Here we argue for replications as a core type of class assignment in archaeology courses to close this gap and establish a culture of replication and reproducibility. We review replication assignments in other fields, and describe how to implement a replication assignment suitable for many types of archaeology programs. We describe our experience with replication in an upper level undergraduate class on stone artefact analysis. Replication assignments can help archaeology programs give students skills that enable transparent and reproducible research.  
---

@@

In his influential study of replication, sociologist of science Harry Collins argued that replication is at the core of scientific practice, writing “Replicability [...] is the Supreme Court of the scientific system” [@collins1992changing p. 19]. Like other observers of science, Collins claimed that in replication, private observations become communal facts, offering vital protection from error and fraud. In this paper we propose a new type of assignment for the archaeology classroom, the replication report, to better align the practice of teaching archaeology with the transparency and openness ideals of science [@nosek2015promoting]. The replication report assignment involves three key steps: 1) students analysing a published report to determine the main claims made by the authors of that report, 2) students obtaining the data used by the authors, and 3) analysing that data to determine if one or more of the authors’ claims are reliable.

@@

@barba2018terminologies points out that although there has been prolific discussion of the terms "reproducibility" and "replication" in many disciplines in recent years, confusion and conflicting uses are widespread. In her survey of relevant literature, Barba finds that some fields make no distinction between "reproducibility" and "replication", and among fields that do recognise a distinction between the two, the meanings are sometimes directly inverted. Here we follow what Barba has identified as the most common, long-established and highly cited definitions of these terms, as also recently recommended by the National Academies of Sciences, Engineering, and Medicine (2019). Reproducibility is the ability to obtain results by using the same data, code, and procedures provided by the original authors [@Marwick2017]. This is only possible when the authors make all those materials available, for example in a research compendium [@marwick2018packaging]. Replication is the ability to arrive at the same scientific conclusions in a new study, collecting new data (possibly with different methods) and completing new analyses.

@@

The Freie Universität Berlin extracurricular graduate seminar course 'Digital Open Science' aims to teach open science practice and assigns replication projects, mostly around neuroimaging topics [@toelch2018digital]. These projects are carried out with a variety of typical open science software tools and services, including Python, R, Git, GitHub and the Open Science Framework. Students first receive extensive lectures and hands-on tutoring, then choose a simple neuroscience experiment to replicate. Finally, they present their results at a symposium. The course’s primary goal is to teach students the value of verifying data upon which their own future research might rely. Students report a high rate of engaging in open science practices after taking the class, and 80% of the participants said that they believe the open science techniques will improve their future research as professionals [@toelch2018digital]. Millman et al (2018) describe a similar course at Berkeley that teaches students how to use open science tools to complete a capstone replication assignment on neuroscience topics.

This brief survey demonstrates that replication assignments are widely known in economics, psychology, political science, neuroscience and other fields [e.g. @roettger2018towards] . Common elements include group work, use of open science software and services to make the replication results openly accessible to anyone, and a scaffolded, stepwise approach to the task to ensure that students receive instructor support at multiple stages in the assignment. To the best of our knowledge replication assignments are not common in archaeology programs. We posted a message to the Society of American Archaeology Teaching Archaeology Interest Group e-community on 27 May 2019 to ask for examples of replication assignments used in teaching archaeology, and received no replies from anyone teaching with replication.

@@

In this section we describe our replication assignment and how we assessed its effectiveness. A brief discussion of our replication report assignment was announced at the beginning of the ten week-long term to give students background about the purpose and concepts of replication and our expectations. Our replication report assignment consisted of three small, graded activities to scaffold the preparation of the final report. The first step started from Week Four and each step was separated by one week to give students time to work and submit their final reports due on Week Seven. Students were expected to work in groups of 3-4 people, but submit their assignments for each of the three steps and the final report individually. Submissions for each step were graded as complete/incomplete, with feedback provided individually via the Canvas learning management system, and collectively during class meetings. Our course had no prerequisites, so we assumed no prior knowledge of the R programming language among the students, and were prepared to teach them as complete novices. We chose R [@R_core_team] because it is widely used by archaeologists [@Schmidt2019], and also commonly taught in undergraduate classes in social sciences and statistics [@baumer2014r; @ccetinkaya2018infrastructure; @dvorak2019efficient]. We were also prepared for students to have no prior experience with replication assignments.

@@

Step three of the assignment required students to create a file structure on their computer to organise their assignment files, following basic guidance in @marwick2018packaging. They also had to download an R Markdown file and write a small amount of R code to read in the raw data and explore it with one basic visualisation, using data in the target article. R Markdown is a file format for making reproducible documents with R. An R Markdown document is written in markdown (a simple plain text format) and contains chunks of embedded R code [@xie2018r]. The document can be easily converted into many standard formats, such as Microsoft Word, PDF and HTML. We prepared an R Markdown template file with some basic headings (following the IMRaD or Introduction-Method-Results-and-Discussion format) and empty code chunks to provide guidance to the students on how many code chunks were expected, and where in the document they should appear. As students wrote their R code and encountered errors, they were encouraged to share screenshots on Slack for the instructors to assist with troubleshooting. After completing this step, the instructor met with each group to review the main claims identified by the students, review the visualisation they had chosen to replicate, and provide guidance on writing the R code to produce the key visualisations.

@@

The final task was for the students to write their report, and submit a reproducible research compendium. This included three files: (1) their R Markdown document, (2)  the raw data file, and (3) the output document (e.g. Microsoft Word document that is produced when they knit the R Markdown file). Our expectation was that we could reproduce any student’s results by running their submitted R Markdown document with the raw data file to produce the Word document they submitted. The final report was graded with a rubric which was presented to the students at the first step to help set expectations about what the final product should look like.

@@

The first step, choosing the target article, revealed the need for some intervention from the instructor to guide students to articles that used relatively simple statistical methods. For example, one group initially chose @Breslawski_Etter_Jorgeson_Boulanger_2018 as their target article, but the key claims in this paper depend on multiple comparisons of multilevel regression models. We explained to the students that if they attempted to replicate a key claim of this paper then they would likely be doing substantially more work than other groups in the class. We invited this group to choose a different target article to ensure a more comparable experience, which they accepted. The statistical backgrounds of our students was highly diverse, so we could not expect students to be very discerning about the statistical complexity of the methods in the articles on the list of potential target articles. As a consequence, we were prepared to intervene to guide their selection of a paper that we could be sure they could successfully replicate, given the time available.

@@

```{r show-plot-feedback, fig.cap = "Results of the anonymous feedback survey on the replication assignment"}
include_graphics(here::here("analysis/figures/feedback_likert_yn_plot.png"))
```

@@

```{r show-plot-feedback-correlations, fig.cap = "(ref:correlation-caption)"}
include_graphics(here::here("analysis/figures/feedback_likert_corr.png"))
```

@@

```{r show-plot-rubric, fig.cap = "Distribution of students' scores across the grading rubric criteria. Each point is one student. Red lines indicate the mean score for all students per criterion."}
include_graphics(here::here("analysis/figures/rubric_score_distribution.png"))
```

@@

The challenges of requiring students to do replication assignments are similar to those faced in many types of archaeology classes with quantitative skills at their core. In our case, the absence of a prerequisite and the high variability of statistical and programming skills meant that some students needed a lot more support than others. This may make replication assignments impractical where instructor-student contact hours are limited. In addition to time, the instructor needs to have a high skill level in quantitative methods to guide students in their engagement with the literature. The instructor will also benefit from a high tolerance for helping students to solve coding problems. To mitigate this in the future we will add a step of student peer review [@wessa2009reproducible] to distribute the feedback task beyond the instructor and give students an opportunity to get assistance from each other more formally.

@@

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

@@

##### pagebreak

>>>>

---
title             : "How to use replication assignments for teaching integrity in empirical archaeology"

author:
  -  Ben Marwick
  -  Liying Wang
  -  Ryan Robinson
  -  Hope Loiselle

abstract: |
  The value of new archaeological knowledge is strongly determined by how credible it is, and a key measure of scientific credibility is how replicable new results are. However, few archaeologists learn the skills necessary to conduct replication as part of their training. This means there is a gap between the ideals of archaeological science, and the skills we teach future researchers. Here we argue for replications as a core type of class assignment in archaeology courses to close this gap and establish a culture of replication and reproducibility. We review replication assignments in other fields, and describe how to implement a replication assignment suitable for many types of archaeology programs. We describe our experience with replication in an upper level undergraduate class on stone artefact analysis. Replication assignments can help archaeology programs give students skills that enable transparent and reproducible research.

bibliography      : "references.bib"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
---

@@

In his influential study of replication, sociologist of science Harry Collins argued that replication is at the core of scientific practice, writing “Replicability [...] is the Supreme Court of the scientific system” [@collins1992changing p. 19]. Like other observers of science, Collins claimed that in replication, private observations become communal facts, offering vital protection from error and fraud. In this paper we propose a new type of assignment for the archaeology classroom, the replication report, to better align the practice of teaching archaeology with the transparency and openness ideals of science [@nosek2015promoting]. The replication report assignment involves three four steps: 1) students analysing a published report to determine the main claims made by the authors of that report, 2) students obtaining the data used by the authors, 3) analysing that data to help determine if one or more of the authors’ claims are reliable, and 4) submitting a research compendium that documents the students' work in a reproducible format that includes the code and data used in the assignment.

@@

@barba2018terminologies points out that although there has been prolific discussion of the terms "reproducibility" and "replication" in many disciplines in recent years, confusion and conflicting uses are widespread. In her survey of relevant literature, Barba finds that some fields make no distinction between "reproducibility" and "replication", and among fields that do recognise a distinction between the two, the meanings are sometimes directly inverted. Here we follow what Barba has identified as the most common, long-established and highly cited definitions of these terms, as also recently recommended by the National Academies of Sciences, Engineering, and Medicine [-@national2019reproducibility]. Reproducibility is the ability to obtain results by using the same data, code, and procedures provided by the original authors [@Marwick2017]. This is only possible when the authors make all those materials available, for example in a research compendium [@marwick2018packaging]. Replication is the ability to arrive at the same scientific conclusions in a new study, collecting new data (possibly with different methods) and completing new analyses.

@@

The Freie Universität Berlin extracurricular graduate seminar course 'Digital Open Science' aims to teach open science practice and assigns replication projects, mostly around neuroimaging topics [@toelch2018digital]. These projects are carried out with a variety of typical open science software tools and services, including Python, R, Git, GitHub and the Open Science Framework. Students first receive extensive lectures and hands-on tutoring, then choose a simple neuroscience experiment to replicate. Finally, they present their results at a symposium. The course’s primary goal is to teach students the value of verifying data upon which their own future research might rely. Students report a high rate of engaging in open science practices after taking the class, and 80% of the participants said that they believe the open science techniques will improve their future research as professionals [@toelch2018digital]. @Millman_Brett_Barnowski_Poline_2018 describe a similar course at Berkeley that teaches students how to use open science tools to complete a capstone replication assignment on neuroscience topics.

This brief survey demonstrates that replication assignments are widely known in economics, psychology, political science, neuroscience and other fields [e.g. @roettger2018towards] . Common elements include group work, use of open source software and services to make the replication results openly accessible to anyone, and a scaffolded, stepwise approach to the task to ensure that students receive instructor support at multiple stages in the assignment. To the best of our knowledge replication assignments are not common in archaeology programs, although the tools and data structures are generally similar to other social sciences. We posted a message to the Society of American Archaeology Teaching Archaeology Interest Group e-community on 27 May 2019 to ask for examples of replication assignments used in teaching archaeology, and received no replies from anyone teaching with replication. More broadly in archaeology replication and reproducibility has received limited, but growing, attention. Elsewhere we have documented recent rapid increases in the number of publications that include code and data to enable readers to reproduce the published results [@Marwick_Schmidt_2019]

@@

In this section we describe our replication assignment and how we assessed its effectiveness. A brief discussion of our replication report assignment was announced at the beginning of the ten week-long term to give students background about the purpose and concepts of replication and our expectations. Our replication report assignment consisted of three small, graded activities to scaffold the preparation of the final report. The first step started from Week Four and each step was separated by one week to give students time to work and submit their final reports due on Week Seven. Students were expected to work in groups of 3-4 people, but submit their assignments for each of the three steps and the final report individually. Submissions for each step were graded as complete/incomplete, with feedback provided individually via the Canvas learning management system, and collectively during class meetings. Our course had no prerequisites, so we assumed no prior knowledge of the free and open source R programming language among the students, and were prepared to teach them as complete novices. We chose R [@R_core_team] because it is widely used by archaeologists [@Schmidt2019], and also commonly taught in undergraduate classes in social sciences and statistics [@baumer2014r; @ccetinkaya2018infrastructure; @dvorak2019efficient]. We were also prepared for students to have no prior experience with replication assignments.

@@

Step three of the assignment required students to create a file structure on their computer to organise their assignment files, following basic guidance in @marwick2018packaging. They also had to download an R Markdown file and write a small amount of R code to read in the raw data and explore it with one basic visualisation, using data in the target article. R Markdown is a file format for making reproducible documents with R. An R Markdown document is written in markdown (a simple plain text format) and contains chunks of embedded R code [@xie2018r]. The document can be easily converted into many standard formats, such as Microsoft Word, PDF and HTML, we give more detail about this in @Marwick2017. We prepared an R Markdown template file with some basic headings (following the IMRaD or Introduction-Method-Results-and-Discussion format) and empty code chunks to provide guidance to the students on how many code chunks were expected, and where in the document they should appear. As students wrote their R code and encountered errors, they were encouraged to share screenshots on Slack for the instructors to assist with troubleshooting. After completing this step, the instructor met with each group to review the main claims identified by the students, review the visualisation they had chosen to replicate, and provide guidance on writing the R code to produce the key visualisations.

@@

The final task was for the students to write their report, and submit a reproducible research compendium. This included three files: (1) their R Markdown document, (2)  the raw data file, and (3) the output document (e.g. Microsoft Word document that is produced when they knit the R Markdown file). The students submitted these materials to Canvas for grading. We did not make the student work public, unlike some of examples described above that deposit student work in public on the Open Science Framework. Our expectation was that we could reproduce any student’s results by running their submitted R Markdown document with the raw data file to produce the Word document they submitted. The final report was graded with a rubric which was presented to the students at the first step to help set expectations about what the final product should look like.

@@

The first step, choosing the target article, revealed the need for some intervention from the instructor to guide students to articles that used relatively simple statistical methods. For example, one group initially chose @Breslawski_Etter_Jorgeson_Boulanger_2018 as their target article, but the key claims in this paper depend on multiple comparisons of multilevel regression models. We explained to the students that if they attempted to replicate a key claim of this paper then they would likely be doing substantially more work than other groups in the class. We invited this group to choose a different target article to ensure a more comparable experience, which they accepted. The statistical backgrounds of our students was highly diverse, so we could not expect students to be very discerning about the statistical complexity of the methods in the articles on the list of potential target articles. As a consequence, we were prepared to intervene to guide their selection of a paper that we could be sure they could successfully replicate, given the time available. The target articles used by this class were @Marwick_Clarkson_O_Connor_Collins_2016, @Bicho2020, and @Marwick_2013.

@@

@@

@@

@@

The challenges of requiring students to do replication assignments are similar to those faced in many types of archaeology classes with quantitative skills at their core. In our case, the absence of a prerequisite and the high variability of statistical and programming skills meant that some students needed a lot more support than others. This may make replication assignments impractical where instructor-student contact hours are limited. In addition to time, the instructor needs to have a high skill level in quantitative methods to guide students in their engagement with the literature. The instructor will also benefit from a high tolerance for helping students to solve coding problems, and having a teaching assistent with a suitable background and similar qualities. To mitigate this in the future we will add a step of student peer review [@wessa2009reproducible] to distribute the feedback task beyond the instructor and give students an opportunity to get assistance from each other more formally.

@@

@@

```{r show-plot-feedback, fig.cap = "Results of the anonymous feedback survey on the replication assignment", fig.align='center'}
include_graphics(here::here("analysis/figures/feedback_likert_yn_plot.png"))
```

```{r show-plot-feedback-correlations, fig.cap = "(ref:correlation-caption)", fig.align='center'}
include_graphics(here::here("analysis/figures/feedback_likert_corr.png"))
```

```{r show-plot-rubric, fig.cap = "Distribution of students' scores across the grading rubric criteria. Each point is one student. Red lines indicate the mean score for all students per criterion.", fig.align='center'}
include_graphics(here::here("analysis/figures/rubric_score_distribution.png"))
```
